Multi-State Temporal Occupancy Modeling

### 2020-07-22 Model Statement/MSTOM Statement.tex
MSTOM pdf update

\begin{document}
\maketitle

@@

\subsubsection{Multinomial-logit Model 1}
To model site-level occurrence according to a single continuous covariates ($\mathbf{x}$) that effects state-level probabilities differently, we could specify this as,
\begin{center}
\begin{align*}
\textbf{z}_{i} &\sim \text{Categorical}(\boldsymbol{\psi_{i}})\\
\boldsymbol{\psi}_{i} &= \begin{bmatrix} \psi^1_{i} & \psi^2_{i} & \psi^3_{i} & \psi^4_{i} \end{bmatrix}.\\
\psi^1_{i} &=\frac{\phi^1_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^2_{i} &=\frac{\phi^2_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^3_{i} &=\frac{\phi^3_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^4_{i} &=\frac{\phi^4_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}.\\
\phi^1_{i} &= 1\\
\phi^2_{i} &= e^{\alpha_{1}+\alpha_{2}\times x_{i}}\\
\phi^3_{i} &= e^{\alpha_{3}+\alpha_{4}\times x_{i}}\\
\phi^4_{i} &= e^{\alpha_{5}+\alpha_{6}\times x_{i}}\\
\end{align*}
\end{center}

@@

\subsubsection{Multinomial-logit Model 2}
We could generalize the modeling of state-specific occupancy parameters using separate design matrices ($\textbf{x}_{i}^{\text{Day}}, \textbf{x}_{i}^{\text{Night}}, \textbf{x}_{i}^{\text{ND}}$) that for each site $i$ are  1 x Q$_{m}$ (being the number of columns) and associated vectors of coefficients ($\boldsymbol{\alpha}^{\text{Day}}, \boldsymbol{\alpha}^{\text{Night}}, \boldsymbol{\alpha}^{\text{ND}}$) that are $Q_{m}$ x 1, and specified as,
\begin{center}
\begin{align*}
\textbf{z}_{i} &\sim \text{Categorical}(\boldsymbol{\psi_{i}})\\
\boldsymbol{\psi}_{i} &= \begin{bmatrix} \psi^1_{i} & \psi^2_{i} & \psi^3_{i} & \psi^4_{i} \end{bmatrix}.\\
\psi^1_{i} &=\frac{\phi^1_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^2_{i} &=\frac{\phi^2_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^3_{i} &=\frac{\phi^3_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^4_{i} &=\frac{\phi^4_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}.\\
\phi^1_{i} &= 1\\
\phi^2_{i} &= e^{\textbf{x}_{i}^{\text{Day}}\boldsymbol{\alpha}^{\text{Day}}}\\
\phi^3_{i} &= e^{\textbf{x}_{i}^{\text{Night}}\boldsymbol{\alpha}^{\text{Night}}}\\
\phi^4_{i} &= e^{\textbf{x}_{i}^{\text{ND}}\boldsymbol{\alpha}^{\text{ND}}}\\
\end{align*}
\end{center}

@@

>>>>

\begin{document}
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{0pt}
\maketitle

@@

\subsubsection{Multinomial-logit Model 1}
To model site-level occurrence according to a single continuous covariates ($\mathbf{x}$) that effects state-level probabilities differently, we could specify this as,
\begin{center}
\begin{align*}
\textbf{z}_{i} &\sim \text{Categorical}(\boldsymbol{\psi_{i}})\\
\boldsymbol{\psi}_{i} &= \begin{bmatrix} \psi^1_{i} & \psi^2_{i} & \psi^3_{i} & \psi^4_{i} \end{bmatrix}.\\
\psi^1_{i} &=\frac{\phi^1_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^2_{i} &=\frac{\phi^2_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^3_{i} &=\frac{\phi^3_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^4_{i} &=\frac{\phi^4_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}.\\
\end{align*}
\end{center}

\begin{center}
\begin{align*}
\phi^1_{i} &= 1\\
\phi^2_{i} &= e^{\alpha_{1}+\alpha_{2}\times x_{i}}\\
\phi^3_{i} &= e^{\alpha_{3}+\alpha_{4}\times x_{i}}\\
\phi^4_{i} &= e^{\alpha_{5}+\alpha_{6}\times x_{i}}\\
\end{align*}
\end{center}

@@

\subsubsection{Multinomial-logit Model 2}
We could generalize the modeling of state-specific occupancy parameters using separate design matrices ($\textbf{x}_{i}^{\text{Day}}, \textbf{x}_{i}^{\text{Night}}, \textbf{x}_{i}^{\text{ND}}$) that for each site $i$ are  1 x Q$_{m}$ (being the number of columns) and associated vectors of coefficients ($\boldsymbol{\alpha}^{\text{Day}}, \boldsymbol{\alpha}^{\text{Night}}, \boldsymbol{\alpha}^{\text{ND}}$) that are $Q_{m}$ x 1, and specified as,
\begin{center}
\begin{align*}
\textbf{z}_{i} &\sim \text{Categorical}(\boldsymbol{\psi_{i}})\\
\boldsymbol{\psi}_{i} &= \begin{bmatrix} \psi^1_{i} & \psi^2_{i} & \psi^3_{i} & \psi^4_{i} \end{bmatrix}.\\
\psi^1_{i} &=\frac{\phi^1_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^2_{i} &=\frac{\phi^2_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^3_{i} &=\frac{\phi^3_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}\\
\psi^4_{i} &=\frac{\phi^4_{i}}{\phi^1_{i}+ \phi^2_{i}+\phi^3_{i}+\phi^4_{i}}.\\
\end{align*}
\end{center}

\begin{center}
\begin{align*}
\phi^1_{i} &= 1\\
\phi^2_{i} &= e^{\textbf{x}_{i}^{\text{Day}}\boldsymbol{\alpha}^{\text{Day}}}\\
\phi^3_{i} &= e^{\textbf{x}_{i}^{\text{Night}}\boldsymbol{\alpha}^{\text{Night}}}\\
\phi^4_{i} &= e^{\textbf{x}_{i}^{\text{ND}}\boldsymbol{\alpha}^{\text{ND}}}\\
\end{align*}
\end{center}

@@

\subsubsection{Multinomial-logit Model 2}

We could also model detection parameters by site (\textbf{x}) or survey specific (\textbf{V}) covarariates, such that 
\begin{center}
\begin{align*}
\boldsymbol{q}_{ij} &= \begin{bmatrix} 1 & 0 & 0 & 0 \\ 
									1 &  q^{2,2}_{ij} & 0 & 0 \\ 
									1 & 0 & q^{3,3}_{ij} & 0\\
  								       1 & q^{4,2}_{ij} & q^{4,3}_{ij} & q^{4,4}_{ij}
  								      \end{bmatrix}\\
q^{2,2}_{ij} &=e^{\beta^{\text{Day}}_{0} +\beta^{\text{Day}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{3,3}_{ij} &=e^{\beta^{\text{Night}}_{0} +\beta^{\text{Night}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,2}_{ij} &=e^{\beta^{\text{ND2}}_{0} +\beta^{\text{ND2}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,3}_{ij} &=e^{\beta^{\text{ND3}}_{0} +\beta^{\text{ND3}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}\\
q^{4,4}_{ij} &=e^{\beta^{\text{ND4}}_{0} +\beta^{\text{ND4}}_{1}\times x_{i}+\beta^{\text{Day}}_{2}\times v_{ij}}
\end{align*}
\end{center}
and the probability of detection parameters can be derived as, 
\begin{center}
\begin{align*}
p^{\text{Day}}_{ij} &=\frac{q^{2,2}_{ij}}{1+q^{2,2}_{ij}+0+0}\\
p^{\text{Night}}_{ij} &=\frac{q^{3,3}_{ij}}{1+0+q^{3,3}_{ij}+0}\\
p^{\text{ND1}}_{ij} &=\frac{1}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND2}}_{ij} &=\frac{q^{4,2}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND3}}_{ij} &=\frac{q^{4,3}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}\\
p^{\text{ND4}}_{ij} &=\frac{q^{4,4}_{ij}}{1+q^{4,2}_{ij}+q^{4,3}_{ij}+q^{4,4}_{ij}}.
\end{align*}
\end{center}

Diffuse priors for all $\beta$ parameters could be specifed as Logistic(0,1).
